Machine Learning based TIR predictor - Methods

### 2020-09-28 methods.tex
Squashed 'sub/SynBio_ML_Writing/' changes from 15a9eb6..fc97298
fc97298 update plots name
ba377d7 Number of changes to text flow.
639b4df change plots
0dc59c9 revise ML part as EE and MH suggested.
f2eaa9a update WDS kernel description; add exp-exp explanation.

git-subtree-dir: sub/SynBio_ML_Writing
git-subtree-split: fc97298311174519cf4090111f327499dfded902

\subsection{Plasmid Design}

We have used the pBbB6c-GFP plasmid for all our designs. This plasmid comes with GFP mut3b CDS inducible with addition of IPTG. The original RBS for the GFP CDS was replaced with combination of PCR and isothermal assembly. Primers and the assembly strategy have been generated using the Teselagen DESIGN software (Teselagen Biotechnology).

\subsection{PCR}
PCR amplification of the cloning inserts was done using Q5 High-Fidelity 2X Master Mix (NEB, catalogue no. M0492L). 20 $\mu$L reactions were prepared by dispensing each of the 10 $\mu$M reverse primers into a well of a 96-well PCR plate using the Labcyte Echo Liquid Handler. A mastermix consisting of polymerase premix, plasmid DNA template, and the single 10 forward primer was prepared by and dispensed by hand. Reactions were run using Touchdown PCR or standard PCR cycling methods in BioRad C1000 thermal cyclers. Then, samples were incubated at 37$^{\circ}$C for 60 minutes, followed by a 20-minute heat inactivation step at 80$^{\circ}$C.
Capillary electrophoresis of PCR products was performed using the Agilent Technologies ZAG DNA Analyzer system. 2$\mu$L of each PCR reaction was electrophoresed using the ZAG 130 dsDNA Kit (75-20000bp) or ZAG 110 dsDNA Kit (35-5000bp) (Agilent Technologies, catalogue no. ZAG-110-5000; ZAG-130-5000). ProSize Data Analysis Software (Agilent Technologies) was used to generated gel images from the sample chromatograms and sizes were estimated by reference to the upper and lower DNA markers spiked into each sample and a DNA ladder run in well H12 of each sample plate.

\subsection{Isothermal DNA Assembly}
Constructs were assembled using NEBuilder HiFi DNA Assembly Master Mix (NEB, catalogue no. E2621L). Samples were incubated at 37$^{\circ}$C for 60 minutes, followed by a 20-minute heat inactivation step at 80$^{\circ}$C. Reactions consisting of the common fragment and the variable fragment were prepared using the Echo acoustic liquid handler, to a final volume of 5 or 10\(\mu\)L . Assemblies were run in the thermal cycler for 1 hour at 50$^{\circ}$C, followed by an infinite hold step at 4$^{\circ}$C.

\subsection{\textit{E. coli} transformation}
The DH5α cell line (Thermo Fisher Scientific, catalogue no. 18265017) was made chemically competent using the Mix $\&$ Go \textit{E. coli} Transformation Kit $\&$ Buffer Set (Zymo Research, catalogue no. T3001). 20$\mu$L of cells was aliquoted into each well of a cold 96-well PCR plate and stored at -80$^{\circ}$C for later use. Plates of cells were thawed on a -20$^{\circ}$C cold block before 3$\mu$L of the assembly product was added and mixed using the CyBio FeliX liquid handler. Cells were incubated on a cold block for 2-5 minutes before being plated in a 96 square grid on Omnitrays containing LB (BD, catalogue no. ***) with 34$\mu$g/mL chloramphenicol (Sigma, catalogue no. ***). Multiple dilutions of cells in LB were prepared and plated in parallel on Omnitrays in 96 square grid. Plates were incubated overnight at 37$^{\circ}$C.

\subsection{Automated colony picking and culturing}
A Singer Instruments PIXL colony picker was used to select individual colonies from the transformation plates. Each selected colony was used to inoculate 1mL of selective medium in a 2mL square well 96 plate. They were then cultured overnight in 37$^{\circ}$C with shaking (~300rpm).

\subsection{Glycerol stock preparation}
100$\mu$L of sterile 80\% (v/v) glycerol and 100$\mu$L of overnight culture were combined in the wells of a 96 deep (2mL) round well plate using the CyBio Felix liquid handler. They were then sealed with a 96-well silicon sealing mat and transferred to a -80$^{\circ}$C freezer.

\subsection{Culture analysis}
Overnight cultures were started by inoculating 1mL of LB medium supplemented with 34$\mu$g/mL chloramphenicol with ~2$\mu$L of the glycerol stock in a 96 deep (2mL) round well plate. Cultures were incubated at 37$^{\circ}$C with shaking (~300rpm) for ~17 hours. The following morning, 20$\mu$L of overnight culture was added to 980$\mu$L of fresh selection medium and precultures were grown at 37$^{\circ}$C with shaking in 2mL round well 96 plate. 
After 90 minutes, two parallel cultures were prepared in flat-bottom clear polystyrene 96-well plates and were induced with 500$\mu$M IPTG – one plate of 300$\mu$L cultures for flow cytometry analysis (induced with 1.5$\mu$L of 0.1M IPTG) and one plate of 200$\mu$L cultures for plate reader analysis (induced with 1.0$\mu$L of 0.1M IPTG).
•	Cytation 5 acquisition and incubation/shaking settings
•	CytoFLEX S acquisition settings

@@

\begin{figure}[t]
\centering
\subfloat{\label{main:c}\includegraphics[scale=.49]{plots/Recommend_2_Data_Points:_Top_2_UCB.pdf}}\par\medskip
\begin{minipage}{.5\linewidth}
\centering
\subfloat[]{\label{main:a}\includegraphics[scale=.49]{plots/GP-BUCB:_Recommend_1st_Data_Point.pdf}}
\end{minipage}%
\begin{minipage}{.5\linewidth}
\centering
\subfloat[]{\label{main:b}\includegraphics[scale=.49]{plots/GP-BUCB:_Recommend_2nd_Data_Point.pdf}}
\end{minipage}
\caption{Batch Recommendation. We use the batch size of 2, with 5 initial observations. The design (recommendation) space is 24 uniformly distributed points in the range [-2,10], i.e. {-2, -1.5, -1, ..., 9.5, 10}.
The confidence interval are shown with predicted mean of $\pm$ 1.96 standard deviation.
(a) Top UCB recommendations. The recommendations are 2 data points with top UCB scores, constructed with GP predictions. 
(b)(c) GP-BUCB recommendations. (b) shows the first recommended sequence, (c) shows the new predicted confidence interval and the second recommendation based on that.}
\label{fig:batch rec}
\end{figure}

A key property of Gaussian Process regression is that the predictive variance  depends only on observed points (i.e. features), but not on the labels of those observed points. 
% We illustrate this fact by comparing the prediction with the true function value of the new recommended point (Figure \ref{fig: Recommend 1 Data Point (New Observation)}) and the prediction with the predicted value of the new recommended point (Figure \ref{fig: Recommend 1 Data Point (Predicted Mean)}).
% As we can see, the predicted confidence interval is the same for the two plots. 
We can make use of this property to design batch upper confidence bound (BUCB) algorithm \cite{desautels2012parallelizing}.
That is, we can recommend sequences sequentially by updating the UCB score with the updated predicted SD with the the previously recommended points.
% Figure \ref{fig: Recommend 2 Data Points: GP-BUCB.} shows an example of batch UCB recommendation. 
% The plot shows the predicted mean and variance after observing 4 data points. 
First, GP-BUCB  recommends the data point with maximum UCB score based on the predictions over initial 5 observations as shown in Figure \ref{fig:batch rec}(b).  
Then we add the recommended data point ($x = 2$) into the training data set with the predicted mean of that point as label (note it is not the true label, i.e. observation), and update the predicted variance and then we finally update the UCB scores. We then recommend the second data point based on the new UCB scores.   
As we can see in Figure \ref{fig:batch rec}(c), since we assume we have observed $x = 2$, then the new predicted variance of the data points in design space around $x =2$  decreases, so instead of recommending a similar data point $x = 2.5$, we recommend the data point which is in a different maximum area with potentially high labels ($x = 7.5$).     
By using GP-BUCB, we can ensure that the sequences with high similarities will not be recommended in the same batch hence increasing the exploration efficiency.

>>>>

\subsection{Laboratory experimental design}

\subsubsection{BUILD: Construction of genetic devices}

\textbf{Plasmid Design.}
We have used the pBbB6c-GFP plasmid for all our designs.
This plasmid comes with GFP mut3b CDS inducible with addition of IPTG.
The original RBS for the GFP CDS was replaced with combination of PCR and isothermal assembly.
Primers and the assembly strategy have been generated using the Teselagen DESIGN software (Teselagen Biotechnology).

\textbf{PCR.}
PCR amplification of the cloning inserts was done using Q5 High-Fidelity 2X Master Mix (NEB, catalogue no. M0492L).
20 $\mu$L reactions were prepared by dispensing each of the 10 $\mu$M reverse primers into a well of a 96-well PCR plate using the Labcyte Echo Liquid Handler.
A mastermix consisting of polymerase premix, plasmid DNA template, and the single 10 forward primer was prepared by and dispensed by hand. Reactions were run using Touchdown PCR or standard PCR cycling methods in BioRad C1000 thermal cyclers.
Then, samples were incubated at 37$^{\circ}$C for 60 minutes, followed by a 20-minute heat inactivation step at 80$^{\circ}$C.
Capillary electrophoresis of PCR products was performed using the Agilent Technologies ZAG DNA Analyzer system.
2$\mu$L of each PCR reaction was electrophoresed using the ZAG 130 dsDNA Kit (75-20000bp) or ZAG 110 dsDNA Kit (35-5000bp) (Agilent Technologies, catalogue no. ZAG-110-5000; ZAG-130-5000).
ProSize Data Analysis Software (Agilent Technologies) was used to generated gel images from the sample chromatograms and sizes were estimated by reference to the upper and lower DNA markers spiked into each sample and a DNA ladder run in well H12 of each sample plate.

\textbf{Isothermal DNA Assembly.}
Constructs were assembled using NEBuilder HiFi DNA Assembly Master Mix (NEB, catalogue no. E2621L).
Samples were incubated at 37$^{\circ}$C for 60 minutes, followed by a 20-minute heat inactivation step at 80$^{\circ}$C.
Reactions consisting of the common fragment and the variable fragment were prepared using the Echo acoustic liquid handler, to a final volume of 5 or 10\(\mu\)L.
Assemblies were run in the thermal cycler for 1 hour at 50$^{\circ}$C, followed by an infinite hold step at 4$^{\circ}$C.

\textbf{\textit{E. coli} transformation.}
The DH5α cell line (Thermo Fisher Scientific, catalogue no. 18265017) was made chemically competent using the Mix $\&$ Go \textit{E. coli} Transformation Kit $\&$ Buffer Set (Zymo Research, catalogue no. T3001).
20$\mu$L of cells was aliquoted into each well of a cold 96-well PCR plate and stored at -80$^{\circ}$C for later use.
Plates of cells were thawed on a -20$^{\circ}$C cold block before 3$\mu$L of the assembly product was added and mixed using the CyBio FeliX liquid handler.
Cells were incubated on a cold block for 2-5 minutes before being plated in a 96 square grid on Omnitrays containing LB (BD, catalogue no. ***) with 34$\mu$g/mL chloramphenicol (Sigma, catalogue no. ***).
Multiple dilutions of cells in LB were prepared and plated in parallel on Omnitrays in 96 square grid. Plates were incubated overnight at 37$^{\circ}$C.

\textbf{Automated colony picking and culturing.}
A Singer Instruments PIXL colony picker was used to select individual colonies from the transformation plates.
Each selected colony was used to inoculate 1mL of selective medium in a 2mL square well 96 plate.
They were then cultured overnight in 37$^{\circ}$C with shaking (~300rpm).

\textbf{Glycerol stock preparation.}
100$\mu$L of sterile 80\% (v/v) glycerol and 100$\mu$L of overnight culture were combined in the wells of a 96 deep (2mL) round well plate using the CyBio Felix liquid handler.
They were then sealed with a 96-well silicon sealing mat and transferred to a -80$^{\circ}$C freezer.

\subsubsection{TEST: Culture analysis}

\textbf{Test strain culture.}
Overnight cultures were started by inoculating 1mL of LB medium supplemented with 34$\mu$g/mL chloramphenicol with ~2$\mu$L of the glycerol stock in a 96 deep (2mL) round well plate.
Cultures were incubated at 37$^{\circ}$C with shaking (~300rpm) for ~17 hours. The following morning, 20$\mu$L of overnight culture was added to 980$\mu$L of fresh selection medium and precultures were grown at 37$^{\circ}$C with shaking in 2mL round well 96 plate. 
After 90 minutes, two parallel cultures were prepared in flat-bottom clear polystyrene 96-well plates and were induced with 500$\mu$M IPTG – one plate of 300$\mu$L cultures for flow cytometry analysis (induced with 1.5$\mu$L of 0.1M IPTG) and one plate of 200$\mu$L cultures for plate reader analysis (induced with 1.0$\mu$L of 0.1M IPTG).

\textbf{Microplate spectrophotometry.}
The plates were tested in Cytation5 microplate reader.
Cytation 5 acquisition and incubation/shaking settings

\textbf{Flow cytometry.}
The plates were tested in Beckman Coulter CytoFLEX S flowcytometer.
CytoFLEX S acquisition settings

@@

\begin{figure}[t]
\centering
\subfloat{\label{main:c}\includegraphics[scale=.4]{plots/Recommend_2_Data_Points:_Top_2_UCB.pdf}}\par\medskip
\begin{minipage}{.5\linewidth}
\centering
\subfloat[]{\label{main:a}\includegraphics[scale=.4]{plots/GP-BUCB_Recommend_1st_Data_Point.pdf}}
\end{minipage}%
\begin{minipage}{.5\linewidth}
\centering
\subfloat[]{\label{main:b}\includegraphics[scale=.4]{plots/GP-BUCB_Recommend_2nd_Data_Point.pdf}}
\end{minipage}
\caption{Batch Recommendation. We use the batch size of 2, with 5 initial observations. The design (recommendation) space is 24 uniformly distributed points in the range [-2,10], i.e. {-2, -1.5, -1, ..., 9.5, 10}.
The confidence interval are shown with predicted mean of $\pm$ 1.96 standard deviation.
(a) Top UCB recommendations. The recommendations are 2 data points with top UCB scores, constructed with GP predictions. 
(b)(c) GP-BUCB recommendations. (b) shows the first recommended sequence, (c) shows the new predicted confidence interval and the second recommendation based on that.}
\label{fig:batch rec}
\end{figure}

A key property of Gaussian Process regression is that the predictive variance  depends only on observed points (i.e. features), but not on the labels of those observed points. 
% We illustrate this fact by comparing the prediction with the true function value of the new recommended point (Figure \ref{fig: Recommend 1 Data Point (New Observation)}) and the prediction with the predicted value of the new recommended point (Figure \ref{fig: Recommend 1 Data Point (Predicted Mean)}).
% As we can see, the predicted confidence interval is the same for the two plots. 
We can make use of this property to design batch upper confidence bound (BUCB) algorithm \cite{desautels2012parallelizing}.
That is, we can recommend sequences sequentially by updating the UCB score with the updated predicted SD with the the previously recommended points.
% Figure \ref{fig: Recommend 2 Data Points: GP-BUCB.} shows an example of batch UCB recommendation. 
% The plot shows the predicted mean and variance after observing 4 data points. 
First, GP-BUCB  recommends the data point with maximum UCB score based on the predictions over initial 5 observations as shown in Figure \ref{fig:batch rec}(b).  
Then we add the recommended data point ($x = 2$) into the training data set with the predicted mean of that point as label (note it is not the true label, i.e. observation), and update the predicted variance and then we finally update the UCB scores. We then recommend the second data point based on the new UCB scores.   
As we can see in Figure \ref{fig:batch rec}(c), since we assume we have observed $x = 2$, then the new predicted variance of the data points in design space around $x =2$  decreases, so instead of recommending a similar data point $x = 2.5$, we recommend the data point which is in a different maximum area with potentially high labels ($x = 8$).     
By using GP-BUCB, we can ensure that the sequences with high similarities will not be recommended in the same batch hence increasing the exploration efficiency.
